Modular Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-12-07 doc/om-paper/paper.tex
wip on ASM section

% ------------------------------------------------------------
\subsection{Abstract State Machines Observers}\label{sec:abstract}\label{sec:asms}
\lee{These are really observers. Let's call them abstract state machine observers and add a bit of the synchronous observers text here. Also, cite disjunctive invariants paper?}

In model-checking, counterexamples to properties are traces of states. A single state includes assignments to each of the variables, and the counterexample traces we examine (see Section~\ref{sec:experimental}) can contain hundreds of state variables. In practice, the essence of a counterexample depends on a small subset of these. The user has to manually separate the state variables that matter from those that do not for a specific property. Moreover, It is not enough to prevent variables that stutter from displaying since some state variables might be updated without affecting the property. The problem is particularly pertinent when formal models have sufficient detail for implementation synthesis, as is the goal in our work.

@@

But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by a state machine imposed over it. The state machine describes discrete modes or configurations. Each mode captures a relation on a subset of the state variables, as well as the set of possible modes that follow.

This idea is an implementation of abstract state machines (ASMs). ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}. In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:

\begin{definition}[Abstract state machine (ASM)]

\lee{Ben will redo. I'd just remove the Gurevich reference}
\ben{I still need to do this!}

@@

\begin{itemize}
\item {\bf $\OMH(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
\item {\bf $\OMH(m)$, $m > 0$}:
  \begin{enumerate}
  \item $g$ broadcasts a value to each lieutenant, $l$.
  \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$. Then for each $l$, execute $\OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants. $l$ sends $l_v$ or $R$, if $l_v = \ERR$.
  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $\OMH(m-1)$. Compute the majority value over the remaining values, or $V$ if there is no majority. If the majority value is $R$, return $E$.
  \end{enumerate}
\end{itemize}

@@

We have implemented $\OMH(1)$ (as well as the variants, described in Section~\ref{sec:modular}) in Symbolic Analysis Laboratory (SAL)~\cite{SAL}. SAL contains a suite of model-checkers. In our work, we use infinite-state (SMT-based) $k$-induction~\cite{Moura_boundedmodel,Armando-et-al}.

We follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: \emph{relays} and \emph{receivers}. Relays \lee{finish}

We refer to the general as the \emph{source}. Thus, for example, in the case where four nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers. We model communication through one-way, typed channels. The source broadcasts a message to each relay which, in turn, each broadcast their messages to all receivers.

>>>>

\subsection{Proof by Abstraction}\label{sec:abstract}\label{sec:asms}
\lee{These are really observers. Let's call them abstract state machine observers and add a bit of the synchronous observers text here. Also, cite disjunctive invariants paper?}

In model-checking, counterexamples to properties are traces of states. A state is an assignment to each of the system variables and the counterexample traces we examine (see Section~\ref{sec:experimental}) can contain hundreds of system  variables. In practice however, the essence of a counterexample depends on a small subset of these. The user has to manually separate the state variables that matter from those that do not for a specific property. Moreover, It is not enough to prevent variables that stutter from displaying since some state variables might be updated without affecting the property. The problem is particularly pertinent when formal models have sufficient detail for implementation synthesis, as is the goal in our work.

@@

But the relationship between many of the state variables often has an intuitive structure, one familiar to the designer, that can be characterized by an auxiliary state machine imposed over it. The state space of this auxiliary machine comprises discrete modes or configurations of the real system. Each mode captures a relation on a subset of the state variables, as well as the set of possible modes that follow.

\cite{timed-systems}
\cite{Rushby00:CAV}

% This idea is an implementation of abstract state machines (ASMs). ASMs are a generalization of finite state machines originally developed by Gurevich to generalize state machines to operate over arbitrary data structures~\cite{asm}. In our modeling framework we take the ASM idea and turn it into both and auxiliary state machine and an observer of the real system as follows.

% In our context, we require only a restricted form of ASMs known as a \emph{sequential small-step ASM}:

\begin{definition}[Abstract state machine observer (ASMO)]

@@

\begin{itemize}
\item {\bf $\OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $\ERR$).
\item {\bf $\OM(m)$, $m > 0$}:
  \begin{enumerate}
  \item $g$ broadcasts a value to each lieutenant, $l$.
  \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$. Then for each $l$, execute $\OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants. $l$ sends $l_v$ or $R$, if $l_v = \ERR$.
  \item For each lieutenant $l \in L$, remove all $\ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$. Compute the majority value over the remaining values, or $V$ if there is no majority. If the majority value is $R$, return $E$.
  \end{enumerate}
\end{itemize}

@@

We have implemented $\OMH(1)$ (as well as many variants, described in Section~\ref{sec:modular}) in the Symbolic Analysis Laboratory~\cite{SAL} (SAL) language. The verification is also carried out in SAL, which contains a suite of model-checkers. In our work, we use infinite-state (SMT-based) $k$-induction~\cite{Moura_boundedmodel,Armando-et-al}.

We have chosen to follow Rushby~\cite{Rushby:OM1} in ``unrolling'' the communication among lieutenants into two sets of logical nodes: relays and receivers. A special node called the source takes the place of the general above. Thus, for example, in the case where 4 nodes execute $\OMH(1)$, our model consists of one source, three relays, and three receivers. We model communication through one-way, typed channels. The source broadcasts a message to each relay which, in turn, each broadcast their messages to all receivers.
